Robust Estimation, Prediction and Control with \\ Linear Dynamics and Generic Costs

### 2020-02-05 main.tex
Update paper

Despite the dazzling successes of Reinforcement learning in cracking many problems considered unsolvable in the past decade [ref needed], this tool has hardly been applied in real industrial issues. This could be attributed to two undesirable properties which limit its practical applications: the dependence on a tremendous amount of interaction data that cannot always be simulated, and the reliance on trial-and-error and random exploration. In this paper, we consider the problem of controlling an unknown linear system $x(t)$ so as to maximise an \emph{arbitrary} bounded reward function $R$, in a critical setting where mistakes are costly and must be avoided at all time. 
This choice of rich reward space is crucial to have sufficient flexibility to model non-convex and non-smooth functions that naturally arise in many practical problems such as combinatorial optimisation, obstacle avoidance, etc., while quadratic costs are only suited for the tracking of a reference trajectory known a priori.

@@

The dynamics of the real system are described in continuous time. However, all the sensing and control performed on-board happen in discrete time, with time-step $\dd t$. For any variable $z$, we use the subscript notation to refer to these discrete times: $z_n = z(t_n)$ with $t_n = n\dd t$. We use bold symbols to denote temporal sequences $\bz = (z_n)_{n\in\Natural}$.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=0.6\linewidth]{img/ellipsoid}
	\caption{The model estimation procedure. The size of the confidence region $C_{[N],\delta}$ shrinks with the number of samples $N$.}
\end{figure}

@@

At the time step $n=N$, considering the confidence region $\cC_{[N],\delta}$ obtained in the model estimation phase, we formulate our robust control objective:
\begin{equation}
\label{eq:robust-control}
\sup_{\bu\in{(\Real^q)}^\Natural} \underbrace{\inf_{\substack{A(\theta) \in \cC_{[N],\delta} \\ \bom\in[\underline{\bom},\overline{\bom}]^\Natural}} \left[\sum_{n=N+1}^\infty \gamma^n R(x_n(\bu,\bom))\right]}_{V^r(\bu)}
\end{equation}
where $\bu = (u_{N+1},u_{N+2},\dots)$ and $\bom = (\omega_N, \omega_{N+1},\dots)$ are the sequences of future controls and perturbations, respectively. The rest of the paper will tackle the optimisation of this objective.

@@

Minimax problems such as \eqref{eq:robust-control} are notoriously difficult to solve even in cases such as Linear-Quadratic systems (LQ) where the reward $R$ has a simple form. Here with an arbitrary function $R$, we cannot hope to derive a solution in explicit form. Our second contribution is to propose to solve \eqref{eq:robust-control} numerically by adopting a \emph{predict-then-optimise} scheme. Thus, in \autoref{sec:prediction}, we aim to derive a \emph{set predictor} $X(t)$ for the system \eqref{eq:dynamics}, which takes the information on the observed current state ${x}_N$, a confidence region $\cC_\delta$ verifying \eqref{eq:confidence}, a planned control sequence $\bu$ and some admissible bounds on the perturbation $[\underline{\omega}(t),\overline{\omega}(t)]$; and produces predictions of sets $X(t)$ that verify the inclusion property:
\begin{equation}
 x(t)\in X(t),\quad\forall t\geq t_N.
\end{equation}

@@

\paragraph{Linear Quadratic systems} There is a rich body of literature dedicated to the study of LQ systems, and we focus on those providing non-asymptotic results when the system parameters $A, B$ are unknown. In their seminal work \citet{abbasi-yadkori11a}, consider the problem of cumulative regret minimisation. They follow the \emph{Optimism in the Face of Uncertainty} paradigm from the multi-armed bandit literature, that consists in selecting an optimistic system within a high-confidence region while enforcing a controllability constraint, and computing the corresponding optimal control in closed-form by solving a Riccati equation. They show that this procedures achieves a $\tilde{\cO}\left(T^{1/2}\right)$ cumulative regret. \citet{abeille18a} provide the same result by applying Thomson Sampling rather of optimism to select a dynamical model, and rejection sampling to enforce the controllability constraint. These methods aim at regret minimization and explore actively to quickly reduce uncertainty, through information-seeking behaviours. Other work use noise injection for exploration such as \citep{Dean2017} or \citep{Dean2018}. However, being optimistic and minimising a cumulative regret does not fit a critical setting, where ensuring safety is the first criteria requires instead to consider the pessimistic outcomes.
The work of \citet{Dean2017} is closest to our setting: after an offline estimation phase under random exploration, they estimate a simple regret between a robust controller and the oracle optimal performance. 
We also focus on where previous methods often suffer from a limited applicability and computationally intractable subroutines, with the exception of \cite{Dean2018} who provide a polynomial-time algorithm.

@@

\begin{theorem}[Confidence ellipsoid, matrix version of \citep{Abbasi2011}]
\label{thm:confidence_ellipsoid}
Under \autoref{assumpt:gaussian-noise}, it holds with probability at least $1-\delta$ that
\begin{align}
    \label{eq:confidence-ellipsoid}
    \| \theta_{N,\lambda}  - \theta\|_{G_{N,\lambda}} \leq \beta_N(\delta),
\end{align}
with
\begin{equation}
	\label{eq:beta_n}
    \beta_N(\delta)\eqdef \sqrt{2\ln \left(\frac{\det(G_{N,\lambda})^{1/2}}{\delta\det(\lambda I_d)^{1/2}}\right)}
     + (\lambda d)^{1/2}S.
\end{equation}
\end{theorem}

@@

\paragraph{Remark} The robust objective \eqref{eq:robust-control} involves bounds $\underline{\omega}(t)\leq \omega(t) \leq \overline{\omega}(t)$ over the possible perturbations we want to protect against. In the unbounded sub-Gaussian noise setting of \autoref{assumpt:gaussian-noise}, we can use the same formulation by deriving local high-confidence bounds that each holds with probability $\delta_t$. By choosing $\delta_t = \frac{\delta}{t(t+1)}$, we have that the event $\{\forall t, \underline{\omega}(t) \leq \omega(t) \leq \overline{\omega}(t)\}$ holds with probability $1-\delta$.

@@

In order to study the inclusion property \eqref{eq:inclusion-property}, we must start by choosing a set representation $X(t)$. We consider interval prediction, rather than zonotope prediction for instance, for the sake of simplicity of implementation and computational efficiency. Thus, we represent an interval as $X(t) = [\underline{x}(t), \overline{x}(t)]\in \Real^p\times \Real^p$, and the inclusion property becomes:

\begin{equation}
\label{eq:inclusion-property}
\underline{x}(t)\leq x(t)\leq\overline{x}(t),\quad\forall t\geq t_N.
\end{equation}

A simple solution to \eqref{eq:inclusion-property} is proposed in \citep{efimov:hal-00701643}, where, given bounds $\underline{A}\leq A(\theta)\leq\overline{A}$ on $A(\theta)$ from $\cC_{[N],\delta}$ they use matrix interval arithmetics to derive the predictor:
\begin{proposition}
Assuming that \eqref{eq:confidence} is satisfied for the system \eqref{eq:dynamics}, then the interval predictor:
\begin{eqnarray}
\dot{\underline{x}}(t) & = & \underline{A}^{+}\underline{x}^{+}(t)-\overline{A}^{+}\underline{x}^{-}(t)-\underline{A}^{-}\overline{x}^{+}(t) +\overline{A}^{-}\overline{x}^{-}(t)\nonumber \\
 &  &+Bu(t) + D^{+}\underline{\omega}(t)-D^{-}\overline{\omega}(t),\label{eq:predictor-naive}\\
\dot{\overline{x}}(t) & = & \overline{A}^{+}\overline{x}^{+}(t)-\underline{A}^{+}\overline{x}^{-}(t)-\overline{A}^{-}\underline{x}^{+}(t)+\underline{A}^{-}\underline{x}^{-}(t)\nonumber \\
 &  &+Bu(t) + D^{+}\overline{\omega}(t)-D^{-}\underline{\omega}(t),\nonumber \\
 &  & \underline{x}(t_N)=\underline{x}_{t_N},\;\overline{x}(t_N)=\overline{x}_{t_N},\nonumber 
\end{eqnarray}
ensures the inclusion property \eqref{eq:inclusion-property}.
\end{proposition}

However, \citet{leurent2019interval} showed that this predictor can have unstable dynamics, even for stable systems, which can cause a rapid build-up of uncertainty. They propose an enhanced predictor which produces tighter and more stable predictions, and whose stability can be assessed by solving a linear matrix inequality, at the price of an additional requirement:

\begin{assumption}
\label{assumpt:metzler}
There exists an orthogonal matrix $P\in\Real^{d\times d}$ such that $P^\transp A_0 P$ is Metzler\footnote{A matrix is said to be Metlzer when all its non-diagonal coefficients are non-negative.}.
\end{assumption}
In practice, this assumption is often verified. It is for instance the case whenever $A_0$ is diagonalisable, and a fortiori when it is stabilisable. The similarity transformation of \citep{Efimov_a2013} provides a method to compute such $P$ whenever there exist $C_1,C_2$ which make $A(\theta, C_1)$ and $(A_0, C_2)$ observable.
% Under this assumption, without loss of generality up to a change of basis $x\rightarrow P^{-1}x$ that we do not write for the sake of readability, this assumption is equivalent to considering that $A_0$ itself is Metzler.

\begin{theorem}[Interval Predictor \citep{leurent2019interval}]
\label{thm:predictor}
Assuming that \eqref{eq:polytope} and \autoref{assumpt:metzler} are satisfied for the system \eqref{eq:dynamics}, then the interval predictor:
\begin{eqnarray}
\dot{\underline{x}}(t) & = & A_{0}\underline{x}(t)-\Delta A_{+}\underline{x}^{-}(t)-\Delta A_{-}\overline{x}^{+}(t)\nonumber \\
 &  & +Bu(t)+D^{+}\underline{\omega}(t)-D^{-}\overline{\omega}(t),\nonumber\\
\dot{\overline{x}}(t) & = & A_{0}\overline{x}(t)+\Delta A_{+}\overline{x}^{+}(t)+\Delta A_{-}\underline{x}^{-}(t) \label{eq:interval-predictor} \\
 &  & +Bu(t)+D^{+}\overline{\omega}(t)-D^{-}\underline{\omega}(t),\nonumber \\
 &  & \underline{x}(t_N)=\overline{x}(t_N)={x}(t_N)\nonumber 
\end{eqnarray}
ensures the inclusion property \eqref{eq:inclusion-property}.
\end{theorem}

\begin{figure}[tp]
	\centering
	\includegraphics[width=\linewidth]{img/interval-predictor}
	\caption{Comparison of \eqref{eq:predictor-naive} to \eqref{eq:interval-predictor} for a simple system $\dot{x}(t)=-\theta x(t)+d(t)$, with $\theta\in[1, 2]$.}
	\label{fig:predictor_example}
\end{figure}

@@

\label{sec:control}
We approximate the robust objective $V^r$ in \eqref{eq:robust-control} by considering its interval prediction $[\underline{x}(t), \overline{x}(t)]$ based on the dynamics confidence region $\cC_\delta$.

@@

This does not simply amounts to a change of reward function, since the worst case is assessed over the whole past trajectory being optimized, which means that this pessimistic reward $\underline{R_n}$ is not \emph{markovian}.

@@

The consequence of \autoref{prop:lower-bound} is that since all our approximations were conservative, if we manage to find a control sequence such that no \textit{``bad event''} (e.g. a collision) happens according to the surrogate objective $\hat{V}^r$, then we are guaranteed that they will not happen either when the controls are executed on the true system.

\paragraph{Planning}
In order to optimise $\hat{V}^r$, we turn to sequential planning algorithms. We cannot use Dynamic Programming algorithms because the state space is continuous and the pessimistic rewards are non-markovian. Instead, we turn to tree-based planning algorithms which optimise a sequence of actions based on the corresponding sequence of rewards. These methods no dot require the enumeration of the states nor the markovity of the rewards. In particular, we consider the \emph{Optimistic Planning of Deterministic Systems} (\texttt{OPD}) algorithm \citep{Hren2008} tailored for the case where the sequence of rewards is a deterministic function of the sequence of actions. Indeed, the stochasticity of the perturbations and measurements is encased in $\hat{V}^r$: both the predictor dynamics \eqref{eq:interval-predictor} and the corresponding pessimistic rewards \eqref{eq:pessimistic-rewards} are deterministic.

At each planning iteration $k\in[K]$, \texttt{OPD} progressively builds a tree $\cT_{k+1}$ by forming upper-bounds $B_a(k)$ over the sequence values called B-values, and expanding\footnote{The expansion of a leaf node $a$ refers to the simulation of its children transitions $aA = \{ab, b\in A\}$} the leaf $a_k$ with highest B-value:
\begin{equation}
\label{eq:opd}
a_k = \argmax_{a\in\cL_k} B_a(k), \quad B_a(k) = \sum_{n=0}^{h(a)-1} \underline{R}_n(a) + \frac{\gamma^{h(a)}}{1-\gamma}
\end{equation}
where $\cL_k$ is the set of leaves of $\cT_k$, $h(a)$ is the length of the sequence $a$, and $\underline{R}_n(a)$ the pessimistic reward \eqref{eq:pessimistic-rewards} obtained at time $n$ by following the sequence $a$.
\begin{theorem}[Planning performance, \citep{Hren2008}]
\label{theorem:opd-regret}
The simple regret of the \texttt{OPD} algorithm \eqref{eq:opd} applied to the surrogate objective \eqref{eq:surrogate_objective} after $K$ iterations is:
\begin{align*}
\text{if } \kappa>1,\quad& 
\hat{V}^r(a^*) - \hat{V}^r({a_K}) = \cO\left(K^{-\frac{\log 1/\gamma}{\log \kappa}}\right);\\
\text{if }\kappa=1,\quad&
\hat{V}^r(a^*) - \hat{V}^r({a_K}) = \cO\left(\gamma^{\frac{(1-\gamma)^\beta}{c}K}\right)
\end{align*}
where $\kappa$ is a problem-dependent measure of the proportion of near-optimal paths:
\[
\kappa = \limsup_{h\rightarrow\infty} \left|\left\{a\in A^h: \hat{V}^r(a)\geq \hat{V}^r - \frac{\gamma^{h+1}}{1-\gamma}\right\}\right|^{1/h}.
\]
\end{theorem}

Hence, by using enough computational budget $K$ for planning we can get as close as we want to the optimal surrogate value $\hat{V}^r(\bu^*)$, at a polynomial rate. However, there exists a gap between $\hat{V}^r$ and the true robust objective $V^r$, which stems from three approximations: (i) the true reachable set was approximated by an enclosing interval in \eqref{eq:inclusion-property}; (ii) the time-invariance of the dynamics uncertainty $A(\theta)\in\cC_{[N],\delta}$ was handled by the interval predictor \eqref{eq:interval-predictor} as if it were a time-varying uncertainty $\forall t, A(\theta(t))\in\cC_{[N],\delta}$ ; and (iii) the lower-bound $\sum\min\leq \min\sum$ used to define the surrogate objective \eqref{eq:surrogate_objective} is not tight. However, this gap can be bounded with additional assumptions:
\begin{theorem}[Sample complexity]
\label{thm:control-error}
Under some assumptions specified in the proof, which amounts to
\begin{enumerate}
	\item a regularity (lipschitz) assumption for the reward $R$;
	\item a persistent excitation (PE) assumption;
	\item a stability condition in the form of an LMI;
\end{enumerate}
 we can bound the approximation error of \autoref{alg:full}:
\begin{equation*}
\hat{V}^r(\bu) \leq {V}^r(\bu) \leq V(\bu) \leq \hat{V}^r(\bu) + \Delta_\omega + \cO\left({\log N / N}\right)
\end{equation*}
where $\Delta_\omega$ is a constant specified in the proof, which corresponds to an irreductible suboptimality suffered from being robust to instantaneous perturbations $\omega(t)$, with knowledge of $A(\theta)$. % $\max_a \sum_n \min_\omega$
\end{theorem}

@@

The procedure we developed in sections \ref{sec:estimation}, \ref{sec:prediction} and \ref{sec:control} relies on restrictive modelling assumptions, such as the linear dynamics \eqref{eq:dynamics} or \autoref{assumpt:structure}. But what if they are wrong?

@@

Given $N-1$ observations, the \autoref{sec:estimation} provides a polytopic confidence region \eqref{eq:polytope} that contains $A(\theta)$ with probability at least $1-\delta$. Since the dynamics are linear, we can propagate this confidence region to the next observation: $y_{N}$ must belong to the Minkowski sum of a polytope representing model uncertainty $\cP(A_{0} x_N + Bu_N, \Delta A_{1}x_N,\dots, \Delta A_{2^d}x_N)$ and a polytope $\cP(0_p, \underline{\eta}, \overline{\eta})$ bounding the perturbation and measurement noises. \citet{delos2015} provide a way to test this membership in polynomial time using linear programming.

Whenever this membership property is not verified, we can confidently reject the $(A,\phi)$-modelling \autoref{assumpt:structure}. This enables us to consider a rich set of potential features $\left((A^1, \phi^1), \dots, (A^M, \phi^M)\right)$ rather than relying on a single assumption.

@@

We temporarily ignore the parametric uncertainty over $\theta$ to simply consider several known candidate dynamics models, which all correspond to different modelling assumptions.

>>>>

Despite the recent successes of Reinforcement learning in cracking many hard problems, this tool has hardly been applied in real industrial issues. This could be attributed to two undesirable properties which limit its practical applications: the dependence on a tremendous amount of interaction data that cannot always be simulated, and the reliance on trial-and-error and random exploration. In this paper, we consider the problem of controlling an unknown linear system $x(t)$ so as to maximise an \emph{arbitrary} bounded reward function $R$, in a critical setting where mistakes are costly and must be avoided at all time. 
This choice of rich reward space is crucial to have sufficient flexibility to model non-convex and non-smooth functions that naturally arise in many practical problems such as combinatorial optimisation, obstacle avoidance, etc., while quadratic costs are only suited for the tracking of a reference trajectory known a priori.

@@

The dynamics of the real system are described in continuous time. However, all the sensing and control performed on-board happen in discrete time, with time-step $\dd t$. For any variable $z$, we use the subscript notation to refer to these discrete times: $z_n = z(t_n)$ with $t_n = n\dd t$. We use bold symbols to denote temporal sequences $\bz = (z_n)_{n>N}$. We denote $z^+ = \max(z,0)$ and $z^- = z^+-z$, and $[n]$ denotes the interval $\{1,2,\dots, n\}$.

@@

\begin{figure}[t]
	\centering
	\includegraphics[trim={0 2cm 0 0}, clip, width=0.6\linewidth]{img/ellipsoid}
	\caption{The model estimation procedure. The size of the confidence region $C_{[N],\delta}$ shrinks with the number of samples $N$.}
\end{figure}

@@

At the time step $n=N$, considering the confidence region $\cC_{[N],\delta}$ obtained in the model estimation phase, we formulate our robust control objective:
\begin{equation}
\label{eq:robust-control}
\sup_{\bu\in{(\Real^q)}^\Natural} \underbrace{\inf_{\substack{A(\theta) \in \cC_{[N],\delta} \\ \bom\in[\underline{\bom},\overline{\bom}]^\Natural}} \left[\sum_{n=N+1}^\infty \gamma^n R(x_n(\bu,\bom))\right]}_{V^r(\bu)}
\end{equation}
The rest of the paper tackles the optimisation of \eqref{eq:robust-control}.

@@

Minimax problems such as \eqref{eq:robust-control} are notoriously difficult to solve even in cases such as linear-quadratic systems (LQ) where the reward $R$ has a simple form. Here with an arbitrary function $R$, we cannot hope to derive a solution in explicit form. Our second contribution is to propose to solve \eqref{eq:robust-control} numerically by adopting a \emph{predict-then-optimise} scheme. Thus, in \autoref{sec:prediction}, we aim to derive a \emph{set predictor} $X(t)$ for the system \eqref{eq:dynamics}, which takes the information on the observed current state ${x}_N$, a confidence region $\cC_\delta$ verifying \eqref{eq:confidence}, a planned control sequence $\bu$ and some admissible bounds on the perturbation $[\underline{\omega}(t),\overline{\omega}(t)]$; and produces predictions of sets $X(t)$ that verify the inclusion property:
\begin{equation}
 x(t)\in X(t),\quad\forall t\geq t_N.
\end{equation}

@@

\paragraph{Linear-quadratic systems} There is a rich body of literature dedicated to the study of LQ systems, and we focus on those providing non-asymptotic results when the system parameters $A, B$ are unknown. In their seminal work \citet{abbasi-yadkori11a}, consider the problem of cumulative regret minimisation. They follow the \emph{Optimism in the Face of Uncertainty} paradigm from the multi-armed bandit literature, that consists in selecting an optimistic system within a high-confidence region while enforcing a controllability constraint, and computing the corresponding optimal control in closed-form by solving a Riccati equation. They show that this procedures achieves a $\tilde{\cO}\left(T^{1/2}\right)$ cumulative regret. \citet{abeille18a} provide the same result by applying Thomson Sampling rather of optimism to select a dynamical model, and rejection sampling to enforce the controllability constraint. These methods aim at regret minimization and explore actively to quickly reduce uncertainty, through information-seeking behaviours. Other work use noise injection for exploration such as \citep{Dean2017} or \citep{Dean2018}. However, being optimistic and minimising a cumulative regret does not fit a critical setting, where ensuring safety is the first criteria requires instead to consider the pessimistic outcomes.
The work of \citet{Dean2017} is closest to our setting: after an offline estimation phase under random exploration, they estimate a simple regret between a robust controller and the oracle optimal performance. 
We also focus on where previous methods often suffer from a limited applicability and computationally intractable subroutines, with the exception of \cite{Dean2018} who provide a polynomial-time algorithm.

@@

\begin{theorem}[Confidence ellipsoid, matrix version of \citealt{Abbasi2011}]
\label{thm:confidence_ellipsoid}
Under \autoref{assumpt:gaussian-noise}, it holds with probability at least $1-\delta$ that
\begin{align}
    \label{eq:confidence-ellipsoid}
    \| \theta_{N,\lambda}  - \theta\|_{G_{N,\lambda}} \leq \beta_N(\delta),
\end{align}
with
\begin{equation}
	\label{eq:beta_n}
    \beta_N(\delta)\eqdef \sqrt{2\ln \left(\frac{\det(G_{N,\lambda})^{1/2}}{\delta\det(\lambda I_d)^{1/2}}\right)}
     + (\lambda d)^{1/2}S.
\end{equation}
\end{theorem}

@@

\paragraph{Remark} The robust objective \eqref{eq:robust-control} involves bounds $\underline{\omega}(t)\leq \omega(t) \leq \overline{\omega}(t)$ over the possible perturbations we want to protect against. In the unbounded sub-Gaussian noise setting of \autoref{assumpt:gaussian-noise}, we can use the same formulation by deriving local high-confidence bounds that each holds with probability $\delta_n$. By choosing $\delta_n = \frac{\delta}{n(n+1)}$, we have that the event $\{\forall n, \underline{\omega}(t_n) \leq \omega(t_n) \leq \overline{\omega}(t_n)\}$ holds with probability $1-\delta$.

@@

To study the inclusion property \eqref{eq:inclusion-property}, we must start by choosing a set representation $X(t)$. We consider interval prediction, rather than zonotope prediction for instance, for the sake of simplicity of implementation and computational efficiency. Thus, we represent an interval as $X(t) = [\underline{x}(t), \overline{x}(t)]\in \Real^p\times \Real^p$, and the inclusion property becomes:
\begin{equation}
\label{eq:inclusion-property}
\underline{x}(t)\leq x(t)\leq\overline{x}(t),\quad\forall t\geq t_N.
\end{equation}

A simple solution to \eqref{eq:inclusion-property} is proposed in \citep{efimov:hal-00701643}, where, given bounds $\underline{A}\leq A(\theta)\leq\overline{A}$ from $\cC_{[N],\delta}$ they use matrix interval arithmetics to derive the predictor:
\begin{proposition}[Simple predictor of \citealt{efimov:hal-00701643}]
Assuming that \eqref{eq:confidence} is satisfied for the system \eqref{eq:dynamics}, then the interval predictor:
\begin{eqnarray}
\dot{\underline{x}}(t) & = & \underline{A}^{+}\underline{x}^{+}(t)-\overline{A}^{+}\underline{x}^{-}(t)-\underline{A}^{-}\overline{x}^{+}(t) +\overline{A}^{-}\overline{x}^{-}(t)\nonumber \\
 &  &+Bu(t) + D^{+}\underline{\omega}(t)-D^{-}\overline{\omega}(t),\label{eq:predictor-naive}\\
\dot{\overline{x}}(t) & = & \overline{A}^{+}\overline{x}^{+}(t)-\underline{A}^{+}\overline{x}^{-}(t)-\overline{A}^{-}\underline{x}^{+}(t)+\underline{A}^{-}\underline{x}^{-}(t)\nonumber \\
 &  &+Bu(t) + D^{+}\overline{\omega}(t)-D^{-}\underline{\omega}(t),\nonumber \\
 &  & \underline{x}(t_N)=\underline{x}_{t_N},\;\overline{x}(t_N)=\overline{x}_{t_N},\nonumber 
\end{eqnarray}
ensures the inclusion property \eqref{eq:inclusion-property}.
\end{proposition}

However, \citet{leurent2019interval} showed that this predictor can have unstable dynamics, even for stable systems, which causes a fast build-up of uncertainty. They propose an enhanced predictor which exploits the polytopic structure \eqref{eq:polytope} produces tighter and more stable predictions, and whose stability can be assessed by solving an LMI, at the price of an additional requirement:

\begin{assumption}
\label{assumpt:metzler}
There exists an orthogonal matrix $P\in\Real^{d\times d}$ such that $P^\transp A_0 P$ is Metzler\footnote{We say that a matrix is Metlzer when all its non-diagonal coefficients are non-negative.}.
\end{assumption}
In practice, this assumption is often verified. It is for instance the case whenever $A_0$ is diagonalisable, and a fortiori when it is stabilisable. The similarity transformation of \citep{Efimov_a2013} provides a method to compute such $P$. Denote $
\Delta A_{+}=\sum_{i=1}^{2^d}\Delta A_{i}^{+},\;\Delta A_{-}=\sum_{i=1}^{2^d}\Delta A_{i}^{-}$.
% The similarity transformation of \citep{Efimov_a2013} provides a method to compute such $P$ whenever there exist $C_1,C_2$ which make $(A(\theta), C_1)$ and $(A_0, C_2)$ observable.
% Under this assumption, without loss of generality up to a change of basis $x\rightarrow P^{-1}x$ that we do not write for the sake of readability, this assumption is equivalent to considering that $A_0$ itself is Metzler.

\begin{theorem}[Enhanced Predictor of \citealt{leurent2019interval}]
\label{thm:predictor}
Assuming that \eqref{eq:polytope} and \autoref{assumpt:metzler} are satisfied for the system \eqref{eq:dynamics}, then the interval predictor:
\begin{eqnarray}
\dot{\underline{x}}(t) & = & A_{0}\underline{x}(t)-\Delta A_{+}\underline{x}^{-}(t)-\Delta A_{-}\overline{x}^{+}(t)\nonumber \\
 &  & +Bu(t)+D^{+}\underline{\omega}(t)-D^{-}\overline{\omega}(t),\nonumber\\
\dot{\overline{x}}(t) & = & A_{0}\overline{x}(t)+\Delta A_{+}\overline{x}^{+}(t)+\Delta A_{-}\underline{x}^{-}(t) \label{eq:interval-predictor} \\
 &  & +Bu(t)+D^{+}\overline{\omega}(t)-D^{-}\underline{\omega}(t),\nonumber \\
 &  & \underline{x}(t_N)=\overline{x}(t_N)={x}(t_N)\nonumber 
\end{eqnarray}
ensures the inclusion property \eqref{eq:inclusion-property}.
\end{theorem}

\begin{figure}[tp]
	\centering
	{\includegraphics[trim={0 0.6cm 0 0.4cm}, clip, width=\linewidth]{img/interval-predictor}}
    \vspace{-0.5cm}
	\caption{Comparison of \eqref{eq:predictor-naive} to \eqref{eq:interval-predictor} for a simple system $\dot{x}(t)=-\theta x(t)+d(t)$, with $\theta\in[1, 2]$.}
	\label{fig:predictor_example}
\end{figure}

@@

\label{sec:control}
To evaluate the robust objective $V^r$ \eqref{eq:robust-control}, we approximate it thanks to the interval prediction $[\underline{x}(t), \overline{x}(t)]$ of \autoref{sec:prediction}.

@@

This amounts to a change of reward function, except the worst case is assessed over the whole past trajectory being optimized, i.e. this pessimistic reward $\underline{R_n}$ is not \emph{markovian}.

@@

A direct consequence is that since all our approximations were conservative, if we manage to find a control sequence such that no \textit{``bad event''} (e.g. a collision) happens according to the surrogate objective $\hat{V}^r$, then we are guaranteed that they will not happen either when the controls are executed on the true system.

\paragraph{Planning}
To optimise $\hat{V}^r$ \eqref{eq:surrogate_objective}, we cannot use Dynamic Programming algorithms since the state space is continuous and the pessimistic rewards are non-markovian. Instead, we turn to tree-based planning algorithms, which optimise a sequence of actions based on the corresponding sequence of rewards, without requiring markovity nor state enumeration. In particular, we consider the \emph{Optimistic Planning of Deterministic Systems} (\texttt{OPD}) algorithm \citep{Hren2008} tailored for the case when the relationship between actions and rewards is deterministic. Indeed, the stochasticity of the perturbations and measurements is encased in $\hat{V}^r$: both the predictor dynamics \eqref{eq:interval-predictor} and the pessimistic rewards \eqref{eq:pessimistic-rewards} are deterministic.

At each planning iteration $k\in[K]$, \texttt{OPD} progressively builds a tree $\cT_{k+1}$ by forming upper-bounds $B_a(k)$ over the value of sequences of actions $a$, and expanding\footnote{The expansion of a leaf node $a$ refers to the simulation of its children transitions $aA = \{ab, b\in A\}$} the leaf $a_k$ with highest upper-bound:
\begin{equation}
\label{eq:opd}
a_k = \argmax_{a\in\cL_k} B_a(k), \quad B_a(k) = \sum_{n=0}^{h(a)-1} \underline{R}_n(a) + \frac{\gamma^{h(a)}}{1-\gamma}
\end{equation}
where $\cL_k$ is the set of leaves of $\cT_k$, $h(a)$ is the length of the sequence $a$, and $\underline{R}_n(a)$ the pessimistic reward \eqref{eq:pessimistic-rewards} obtained at time $n$ by following the controls $u_n = \pi_{a_n}(x_n)$.
\begin{theorem}[Planning performance, \citep{Hren2008}]
\label{theorem:opd-regret}
The simple regret of the \texttt{OPD} algorithm \eqref{eq:opd} applied to the surrogate objective \eqref{eq:surrogate_objective} after $K$ planning iterations is:
\begin{align*}
\text{if } \kappa>1,\quad& 
\hat{V}^r(a^{\star}) - \hat{V}^r({a_K}) = \cO\left(K^{-\frac{\log 1/\gamma}{\log \kappa}}\right);\\
\text{if }\kappa=1,\quad&
\hat{V}^r(a^{\star}) - \hat{V}^r({a_K}) = \cO\left(\gamma^{{(1-\gamma)^{\log_\gamma (\kappa/|\cA|)}}K/{c}}\right)
\end{align*}
where $\kappa$ is a problem-dependent measure of the proportion of near-optimal paths:
\[
\kappa = \limsup_{h\rightarrow\infty} \left|\left\{a\in A^h: \hat{V}^r(a)\geq \hat{V}^r(a^{\star}) - \frac{\gamma^{h+1}}{1-\gamma}\right\}\right|^{1/h}.
\]
\end{theorem}

Hence, by using enough computational budget $K$ for planning we can get as close as we want to the optimal surrogate value $\hat{V}^r(a^{\star})$, at a polynomial rate. However, there exists a gap between $\hat{V}^r$ and the true robust objective $V^r$, which stems from three approximations: (i) the true reachable set was approximated by an enclosing interval in \eqref{eq:inclusion-property}; (ii) the time-invariance of the dynamics uncertainty $A(\theta)\in\cC_{[N],\delta}$ was handled by the interval predictor \eqref{eq:interval-predictor} as if it were a time-varying uncertainty $A(\theta(t))\in\cC_{[N],\delta},\forall t$ ; and (iii) the lower-bound $\sum\min\leq \min\sum$ used to define the surrogate objective \eqref{eq:surrogate_objective} is not tight. However, this gap can be bounded with additional assumptions:
\begin{theorem}[Sample complexity]
\label{thm:control-error}
Under some assumptions specified in the proof, which amounts to
\begin{enumerate}
	\item a regularity (lipschitz) assumption for the reward $R$;
	\item a persistent excitation (PE) assumption;
	\item a stability condition in the form of an LMI;
\end{enumerate}
 we can bound the approximation error of \autoref{alg:full}:
\begin{equation*}
\hat{V}^r(\bu) \leq {V}^r(\bu) \leq V(\bu) \leq \hat{V}^r(\bu) + \Delta_\omega + \cO\left({\log N / N}\right)
\end{equation*}
where $\Delta_\omega$ is a constant specified in the proof, which corresponds to an irreductible suboptimality suffered from being robust to instantaneous perturbations $\omega(t)$.
\end{theorem}

@@

The procedure we developed in sections \ref{sec:estimation}, \ref{sec:prediction} and \ref{sec:control} relies on strong modelling assumptions, such as the linear dynamics \eqref{eq:dynamics} or \autoref{assumpt:structure}. But what if they are wrong?

@@

Given $N-1$ observations, the \autoref{sec:estimation} provides a polytopic confidence region \eqref{eq:polytope} that contains $A(\theta)$ with probability at least $1-\delta$. Since the dynamics are linear, we can propagate this confidence region to the next observation: $y_{N}$ must belong to the Minkowski sum of a polytope representing model uncertainty $\cP(A_{0} x_N + Bu_N, \Delta A_{1}x_N,\dots, \Delta A_{2^d}x_N)$ and a polytope $\cP(0_p, \underline{\eta}, \overline{\eta})$ bounding the perturbation and measurement noises. \citet{delos2015} provide a way to test this membership in polynomial time using linear programming. Whenever it is not verified, we can confidently reject the $(A,\phi)$-modelling \autoref{assumpt:structure}. This enables us to consider a rich set of potential features $\left((A^1, \phi^1), \dots, (A^M, \phi^M)\right)$ rather than relying on a single assumption, and only retain those that are consistent with the observations so far. Then, every remaining hypothesis must be considered during planning.

@@

We temporarily ignore the parametric uncertainty on $\theta$ to simply consider several candidate dynamics models, which all correspond to different modelling assumptions. We also restrict to deterministic dynamics, which is the case of \eqref{eq:interval-predictor}.

\begin{assumption}[Multi-model ambiguity]
\label{assumpt:multi-model-ambiguity}
The true dynamics $f$ lies within a finite set of candidate models $f^1, \dots, f^M$.
\begin{equation*}
\exists m\in[M]: \dot{x}(t) = f^M(x(t), u(t)), \forall t\geq 0
\end{equation*}
\end{assumption}
We want to adapt our planning algorithm in order to balance these concurrent hypotheses in a robust fashion, i.e. maximise a robust objective with discrete ambiguity:
\begin{equation}
\label{eq:robust-objective-discrete}
V^r = \sup_{a\in\cA^\Natural}\min_{m\in[M]} \sum_{n=N+1}^\infty \gamma^n R_n^m
\end{equation}
where $R_n^m$ is the reward obtain by following the action sequence $a$ up to step $n$ under the dynamics $f^m$.
This objective could be optimised in the same way as in \autoref{sec:control}, but this would result in a coarse and lossy approximation. Instead, we can exploit the finite uncertainty structure of \autoref{assumpt:multi-model-ambiguity} to asymptotically recover the true $V^r$ by modifying the \texttt{OPD} algorithm in the following way:
